---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
  [![R-CMD-check](https://github.com/anjaninandan001/mlr3hf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/anjaninandan001/mlr3hf/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

# mlr3hf

mlr3hf provides a simple interface to download datasets from the Hugging Face Hub and convert them into mlr3 Task objects.

## Installation

You can install the development version of mlr3hf from [GitHub](https://github.com/anjaninandan001/mlr3hf) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("anjaninandan001/mlr3hf")
```

## Example

```{r example, eval = FALSE}
#As of now you can run this to check function is working or not.
library(mlr3hf)
# Download a parquet dataset
result1 <- cached(
  repo_id = "scikit-learn/iris",
  config  = "default",
  parquet = TRUE,
  fun     = download_parquet
)

# Download a CSV dataset
result2 <- cached(
  repo_id   = "scikit-learn/iris",
  file_name = "Iris.csv",
  parquet   = FALSE,
  fun       = download_hfhub
)

print(result1)
print(result2)
```

## Parameters

| Parameter | Description |
|-----------|-------------|
| `repo_id` | Hugging Face dataset repository identifier. |
| `file_name` | Name of the dataset file within the repository. |
| `parquet` | Logical value indicating whether the dataset is in parquet format. |
| `fun` | Function to use for downloading (`download_parquet` or `download_hfhub`). |

## Note

> The package is currently under active development. Some features may be experimental and outputs may change in future releases.