---
title: "Quick revision"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick revision}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(mlr3hf)
```

# Quick Revision

This vignette provides a quick overview of how to use **mlr3hf**.

Suppose you are working on **neuromorphic computing** and need a dataset for your machine learning project.

First, search for relevant datasets on the Hugging Face Hub.

```{r}
library(mlr3hf)
library(mlr3)
list_datasets(search = "neuromorphic computing")
```

This returns a list of dataset repositories matching your search.

Next, create an `HFData` object. At this stage you can inspect the repository before downloading any data.

```{r}
dt <- htsk(repo_id = "Shoriful025/neuromorphic_edge_computing_telemetry")
```

If you want to open the dataset repository in your browser, use:

```{r}
dt$openlink
```

You can also inspect the available dataset configurations and files.

```{r}
dt$configs
dt$siblings
```

Once you know which configuration you want, create the object by specifying the configuration.

```{r}
dt <- htsk(
  repo_id = "Shoriful025/neuromorphic_edge_computing_telemetry",
  config = "default"
)
```

The dataset is downloaded locally, a backend is created, and you can inspect the dataset.

To view the column names or load the data:

```{r}
dt$colnames
dt$data[1:5]
```

You can now perform exploratory data analysis (EDA) using any R package of your choice.

To create an **mlr3** task, specify the target column.

```{r}
task <- as_task(
  dt,
  target_names = "temperature_c"
)
```

If you already know the repository, configuration, and target column, you can create everything directly.

```{r}
dt <- htsk(
  repo_id = "Shoriful025/neuromorphic_edge_computing_telemetry",
  config = "default",
  target = "temperature_c"
)

task <- as_task(dt)
task
```

This concludes a quick revision of the basic **mlr3hf** workflow:

1. Search for datasets.
2. Inspect the repository.
3. Choose a configuration or file.
4. Download the dataset and create a backend.
5. Explore the data.
6. Create an **mlr3** task.
